% This is the acknowledgments page of my

\cleartorecto % A memoir-class command for moving the acknowledgments to a recto page, not verso.
\chapter{Acknowledgements} % For the heading on the page, also registers in the table of contents
\thispagestyle{plain} %This page should have numbers.

First and foremost, I would like to thank my advisor, Elena Aprile, who welcomed me into her research group in 2014.  I owe this dissertation
and my remarkable research experience over the past four years to your guidance and encouragement.  You exemplify scientific excellence and
innovation, and are a role model for young academics.

I would like to give a special thanks to postdoctoral fellows Qing Lin, Patrick de Perio, and Fei Gao.  You committed your time and energy to
help me succeed.  So much of what I have learned over the years is thanks to you.  I feel fortunate to have had you as mentors.

I would also like to thank postdoctoral fellows and research scientists Guillaume Plante, Marc Weber, Luke Goetzke, Marcello Messina, and
Alfio Rizzo.  Your patience and encouragement, especially at the beginning, shaped my four years in the group.

I have had a lot of fun working alongside graduate students Matt Anthony, Yun Zhang, Joey Howlett, and Tianyu Zhu.  The experiments,
analyses, and trips I've shared with you were among the most memorable of my time at Columbia.  I am also happy to have had the
opportunity to know Damon Daw, who worked in our group for a couple of years.

It has been a pleasure knowing Marcela Stern.  You are one of the kindest and most caring people I know, and am grateful for your support
on so many occasions.  It was always a joy talking with you.

I feel extremely fortunate to have gotten to know Jeremy Dodd these last six years.  You are both a mentor and friend, and were there for me
in times of uncertainty.  I am confident that without you I would not be here today.  You had a tremendous impact on my time at Columbia, and
I am forever grateful.

I am especially thankful to my friends and fellow PhD students Felix Clark, Matt Anthony, Laura Havener, Ryne Carbone, Geoff Iwata, and
Russell Smith.  I feel privileged to have made such remarkable, long-lasting friends during this experience.  Your support during
difficult times - both personally and professionally - has been indispensable.  I want to thank Rex Brown, who became a good friend
during his time in New York.  I also want to thank Felix Clark, Rex Brown, and Michael Hammett for being great roommates.

Sarah Goldberg is one of the most remarkable people in my life.  Silly or serious, happy or sad, dumb or intellectual,
being myself around you is so natural.  Your positivity, uniqueness, and kind-heartedness has given our friendship its affluence.

My family has supported and believed in me my entire life.  My parents, Pam and Brock, give me the
confidence and love that motivates me.  My brother, Spencer, is the bravest and most inspiring person I know.  You all have taught me how to
confront adversity, the power of hard work, and importance of joie de vivre.  I am lucky to have such strong relationships with my aunt,
Diana, uncle, Mark, and cousins, Marissa and Tyler, who encourage, love, support me.  I believe people are largely shaped by those closest to
them, so I share this achievement with all of you.  I love you.
